---
title: "Pakistan Rainfall Code"
author: "Rica Rebusit"
date: "5/7/2023"
output:
  html_document:
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Click here to go back to the ANTH333 media presentation: [Media Presentation](https://data485-s22.github.io/website-basic-rrebusit/ANTH333Med.html)

# Load in Necessary Libraries

```{r, warning = FALSE}
library(tidyverse)
```

# Load in Data

```{r}
pakistan <- read.csv("~/Documents/RicaRebusitWeb/Rainfall_1901_2016_PAK.csv")
```

The data is from kaggle and the link to it is right here -\> [Click Me](https://www.kaggle.com/datasets/zusmani/rainfall-in-pakistan)

Information of Data:

This data set has 3 variables:

`Rainfall - (MM)` - This is the recorded rainfall for the year and month at the time. Rainfall is measured in millimeters

`Year` - The year of each recorded rainfall. Contains 116 years of recorded rainfall.

`Month` - The month of each recorded rainfall.

# Exploratory Data Analysis

```{r}
str(pakistan)
```

```{r}
summary(pakistan)
```

```{r}
sum(is.na(pakistan))
```

```{r}
head(pakistan)
```

Renaming `Rainfall - (MM)` to `Rainfall`

```{r}
pakistan <- rename(pakistan, Rainfall = 'Rainfall....MM.')
head(pakistan)
```

Sum the amount of rainfall from all the years and find the percentage rainfall for the month

```{r}
sum_rainfall <- sum(pakistan$Rainfall)
rainfall_percentage <- pakistan %>%
  group_by(Month) %>%
  summarise(Total_1901_2016 = sum(Rainfall), Percent = 100 * Total_1901_2016 / sum_rainfall)
head(rainfall_percentage)
```

Finding the average amount of rainfall

```{r}
average_rainfall <- sum(pakistan$Rainfall)
rainfall_average <- pakistan %>%
  group_by(Month) %>%
  mutate(Average_1901_2016 = mean(Rainfall))
head(rainfall_average)
```

```{r}
ggplot(pakistan, aes(Rainfall)) + geom_histogram(color = "lightblue", fill = "steelblue") + theme_minimal()
```

Saving image for media presentation

```{r}
ggsave("histogram.png")
```

```{r}
ggplot(pakistan, aes(Year, Rainfall, color = Month)) + geom_point(alpha = 0.5) + geom_smooth(se = FALSE)
```

```{r}
ggplot(pakistan, aes(Year, Rainfall, fill = Month)) + geom_bar(stat = "identity") + scale_fill_discrete(name = "Month", limits = month.name)
```

As expected, August and July have higher rainfall due to monsoons

```{r}
ggplot(rainfall_average, aes(Year, Rainfall, color = Month)) + geom_line(alpha = 0.8) + theme_minimal()
```

Saving image for media presentation

```{r}
ggsave("average.png")
```

```{r}
ggplot(subset(pakistan, Month %in% c("June","July", "August")), aes(Year, Rainfall, color = Month)) + geom_line() + facet_wrap(~ Month)
```

Monsoon in 2010

```{r}
monsoon <- filter(pakistan, Month == c("July", "August"), Year == 2010)
monsoon
```

```{r}
ggplot(subset(pakistan, Month %in% c("June","July", "August")), aes(Year, Rainfall, color = Month)) + geom_line() + facet_wrap(~ Month) + geom_point(data = monsoon, aes(Year, Rainfall), color = "black", size = 2) + theme_minimal()
```

Saving image for media presentation

```{r}
ggsave("2010monsoon.png")
```

```{r}
ggplot(subset(pakistan, Month %in% c("March","April", "May")), aes(Year, Rainfall, color = Month)) + geom_line() + facet_wrap(~ Month)
```

```{r}
ggplot(subset(pakistan, Month %in% c("September","October", "November")), aes(Year, Rainfall, color = Month)) + geom_line() + facet_wrap(~ Month)
```

```{r}
ggplot(subset(pakistan, Month %in% c("December","January", "February")), aes(Year, Rainfall, color = Month)) + geom_line() + facet_wrap(~ Month)
```

# Model Choosing

For the sake of brevity, let's use a linear regression model to predict *Rainfall*. As seen above, the data is skewed so we can take log of *Rainfall,* however, implications can come a long with it making the data inaccurate.

```{r}
pakistan$log_rainfall <- log(pakistan$Rainfall)
head(pakistan)
```

```{r}
lm1 <- lm(log_rainfall ~ Year, data = pakistan)
summary(lm1)
```

```{r}
plot(lm1)
```

```{r}
lm2 <- lm(log_rainfall ~ Year + Month, data = pakistan)
summary(lm2)
```

Saving for media presentation

```{r}
save(lm2, file = "lm2.Rdata")
```

```{r}
plot(lm2)
```

Now lets try predicting Rainfall for year 2017 and Month January

Using model 1 first

```{r}
head(predict(lm1))
```

```{r}
logged_prediction <- predict(lm1, data.frame(Year = 2017, Month = "January"))
lm1_prediction <- exp(logged_prediction)
lm1_prediction
```

Now using model 2

```{r}
head(predict(lm2))
```

```{r}
logged_prediction2 <- predict(lm2, data.frame(Year = 2017, Month = "January"))
lm2_prediction <- exp(logged_prediction2)
lm2_prediction
```

We can take a look at accuracy, but for now this ends a quick example of how we can use data science to help predict rainfall.
